

transactions native in Haskell and in Closure


transactions at PL language vs artoicity/ consistnecy / isolation

iolation - opacity of realizartbility vs snapshot islaotion 
		opacity is more than realizability -- aborted transaction are invisible 
		formally written dow semantics of transactions
		
How is isolation achieved
defensively acquire as I go -- when do we acuite locks 
	they all need to keep a log 
			they may nned to abirt when they axceed the cahe
	sometimes implementationtakes care 
	oprimistic - roll forward, they have a readfing log
	pessimistic -- they have an undo log 
		they have a deadlock detection mechanism, and can back-off and restert

sometimes need to acquire the global lock

downside arbirary abirt, interruot

terribe at
	anything with effect; eg output; the oprimisitc version writres at the end
				what if you do not have garbage collection	
version that deletes a neniry location, and another transcation	
		
ordering guarnatees?


Semantics of transacitons
	1) difference between transactionl vs non-transactional acceeees? -- hat are the mixed accesses?
			what isolation do you get? weak or strong isolation?
			in C only non-atomic accesses this creates data-races
	2) nesting 
			no-one 
			isolation -- when does the inner transaction become cislble (flatteing)
			what it the inner birts when the current has succeeded?
			compositionality
			
Not clear what the semantics of nested transactions is!!

when (x){ x:=1; when(y){ y:=1 }}

when(x){ x:= 2l; when(y){ y:= 2 }}

islation is when a write in one transaction becomes visible in the other

what issues would be introduced by our implementation?
	deadlock? avoidance? 
	
Do the semanrics before the impementation

moxed access -- concept does not arise!
	-- make very clear that we expect thar 
	
	
The language exoects that objects obnly whithin a beahviour 

BoC expects that the objects have one owner. -- views?

** OOSPLA 2015 **

atomic --in transaction -- isolation means our atomicity 
all or nithing 

isolation guarantees are impoortnat -- 

isolation -- who sees the effect if a when -- when you spawn a bahaviour 

LEMMA ** If you do not have nete cowsn, thne you have xx serializability guarnatees **

implementation
	GC 
	no llivelock / deadlockk
	side-effects
	
serializability -- we have that --- 