Here we discuss which examples we can and which we cannot prove now

\subsection{Adding external calls to the "safe module's" code}

The method from below can be checked with $Rule_2$. \SP But notice that the spec is too strong! as it assumes that no external entity has access to the password. In that case, nobody can call the method \prg{transfer\_3}

\begin{lstlisting}[language=chainmail, mathescape=true, frame=lines]
class Account

    fld logger: Logger // external class, untrusted

     void transfer_3a (Password p, Account toAcc, int amt) 
     POST:  $\inside{\prg{this.passwd}}$ $\longrightarrow$ [ p==this.password) $\rightarrow$ ...
                                    $\wedge$
                                    $\forall$ a:Account.( a$\neq$fromAcc,toAcc $\rightarrow$ ...    )  ]
     {
          if (p==this.password) 
                  this.balance -= amt
                  toAcc.balance += amt
                  logger.log(this)
     }
 ...
\end{lstlisting}

\SP Should we consider "relative" inside/outside predicates, e.g. $\outsideTo{x} {y}$ holds if we only consider  objects transitively accessible from \prg{y}; that is  $\outsideTo{x} {y}$ holds if none of the external objects transitively accessible from \prg{y} had access to \prg{x}?

Here with the new spec

\begin{lstlisting}[language=chainmail, mathescape=true, frame=lines]
class Account

     field passwd : Password
     field logger: Logger // external class, untrusted

     void transfer_3b (Password p, Account toAcc, int amt) 
     POST: [ $\forall$ a:Account. $\outsideTo{\prg{a.passwd}}{\prg{logger}}$ ] $\longrightarrow$   
                                [ p == this.password) $\rightarrow$ ...
                                  $\wedge$
                                  $\forall$ a:Account.( a$\neq$fromAcc,toAcc $\rightarrow$ ...    )  ]
     {  ... $\mbox{as in }$ transfer_3a  ...  }
 ...
\end{lstlisting}




\subsection{Adding external calls in "our" code --  Me Bean}

We now consider functions which belong  to a class that is not part of $\Mod{_{BA}}$, which use $\Mod{_{BA}}$ as well as some untrusted third party code.  Here we have the module  \prg{MrBean} (ours), which relies on $\Mod{_{BA}}$ in order to pass its account to a third party while preserving some guarantees.

Consider method  \prg{transfer\_4}, which  is similar to  \prg{transfer\_3}. \SP I changed the spec.
\footnote{Here some earlier thoughts on the matter in the new spec: I suppose that we were assuming that we have something like  $\neg \Outside{\prg{a.passwd}}$? Or that ${\prg{a.passwd}}$ is not outside the module 
$\prg{MrBean}\circ\prg{BankAccount}$? Perhaps $\Outside{\_}$ should mention what it is outside of? Perhaps the module \prg{MrBean} would have as invariant that all "its" Accounts keeps their Passwords "inside"?}


\begin{lstlisting}[language=chainmail, mathescape=true, frame=lines]
module MrBean

   class MrAtkinson

      field logger: Logger // external class, untrusted

     void transfer_4 (Account fromAcc, Password p, Account toAcc, int amt) 
     POST:  [  $\forall$ a:Account. $\outsideTo{\prg{a.passwd}}{\prg{logger}}$ ] $\longrightarrow$ 
                                [ p == fromAcc.password) $\rightarrow$ ...
                                  $\wedge$
                                  $\forall$ a:Account.( a$\neq$tfromAcc,toAcc $\rightarrow$ ...    )  ]

     {
          fromAcc.transfer(p,toAcc,amt);
          logger.log(fromAcc)
      }
 ...
\end{lstlisting}


In \prg{transfer\_4} from above, we are passing  an object from
the "safe" module (here \prg{fromAcc}) to the untrusted object. 
What if we were  to pass not only a "safe" object, but also an object from "our" module?
Consider the function \prg{transfer\_5} from below.


\begin{lstlisting}[language=chainmail, mathescape=true, frame=lines]
module MrBean

   class MrAtkinson

    field logger: Logger // external class, untrusted

     // CHALLENGE 
     void transfer_5 (Account fromAcc, Password p, Account toAcc, int amt) 
     POST:  [  $\forall$ a:Account. $\outsideTo{\prg{a.passwd}}{\prg{logger}}$ ] $\longrightarrow$ 
                                [ p == fromAcc.password) $\rightarrow$ ...
                                  $\wedge$
                                  $\forall$ a:Account.  a$\neq$this,toAcc $\rightarrow$ ...     ]

     {   
         fromAcc.transfer(p,toAcc,amt);
         logger.log(fromAcc,this)
      }
 ...
\end{lstlisting}

In \prg{transfer\_5} above, we are passing \prg{this}  to the \prg{logger}. The function \prg{transfer\_5} is
only satisfied if \prg{MrAtkinson} does not leak the password, and does nor make unauthorized payments. So, 
it seems as if proving \prg{transfer\_5} requires a holistic spec for \prg{MrArkinson} too.

\subsection{What can be proven and how}

We can prove  \prg{transfer\_3a} using $Rule_1$, but as noted above, its spec is far too strong.

How could we prove the spec of \prg{transfer\_3a}, and \prg{transfer\_4a} and  \prg{transfer\_4b}.There are the following solutions

\SP The below needs revisiting:

\begin{enumerate}
\item
Include the \prg{Logger}'s code (or its spec) into the proof basis of \prg{Bank}, and make a holistic spec for both of them together. This is a non-solution, because then the logger would not be "external", and "untrusted".
\item
Include the \prg{Logger}'s code (or its spec) into the proof basis of \prg{MrBean}. This is also a non-solution, because then the logger would not be "external", and "untrusted".
\item
Include  \prg{MrBean}'s code  into the proof basis of \prg{Bank}, and make a holistic spec for both of them together. 
\begin{enumerate}
\item
We just prove holistically the "whole" new module. We then only need 
\item
We were hoping that we would be able to "inherit" some of the holistic spec of the \prg{Bank}. As we know, she meaning of $\Outside{\_}$ is delicate as it depends on modules, and appears in both positive and negative positions. But it might be possible.
\end{enumerate}
%\item
%Overwrite the password temporarily before giving it out. Code as in Sect. \ref{s:safe:extern}. 
\item
Introduce the concept of path-accessibility, which might be $Path( {o}, {o'}, {A(o''))}$, and would mean that any path from $o$ to $o'$  must go through an object $o''$ which satisfies  property $A(o'')$
\item
 Adopt $Rule_2$

\end{enumerate}

\noindent,
\textbf{Wrong Conjecture}  Sophia thought that if all the calls in  the body of a function in a client module are either calls to untrusted, or to the safe module (ie no calls to the module at hand), then we should be able to verify it, without holistic spec of the module itself. But \prg{transfer\_5} proves her wrong.

\vspace{.1in}

\textbf{Still Open} How to prove \prg{transfer\_5}  in a modular way?

