\section{Appendix to Section \ref{sect:spec} -- Specifications}
\label{app:spec}

{ \begin{example}[More Method Specifications]
\label{ex:spesMore}
$S_7$ below  guarantees that
\prg{transfer} does not affect the balance of accounts different  from the receiver or argument, and  if the key supplied is not that of the receiver, then no account's balance is affected.  \
$S_8$ guarantees that if the key supplied is that of the receiver, the correct amount is transferred from the receiver to the destination.
 $S_9$ guarantees that \prg{set} preserves the protectedness of a key.

\small{
{\sprepost
		{\strut \ \ \ \ S_7} 
		{ a:\prg{Account}\wedge  a.\prg{\balance}=b \wedge
		(\prg{dst}\neq a\neq\prg{this} \vee \prg{key'}\neq a.\prg{\password})}
	               {\prg{public Account}} {\prg{transfer}} {\prg{dst}:\prg{Account},\prg{key'}:\prg{Key},\prg{amt}:\prg{nat}}
		{ \ a.\prg{\balance}=b\ }
		{  \ a.\prg{\balance}=b\ }
}
\\
{\sprepost 
		{\strut \ \ \ \ \ \ \ \ \ S_8} 
		{  % b,b':\prg{int} \wedge  
		\prg{this}\neq \prg{dst}\wedge \prg{this}.\prg{\balance}=b \wedge  \prg{dst}.\prg{\balance}=b' }
		  {\prg{public Account}}
		  	  {\prg{transfer}} 
		   	 {\prg{dst}:\prg{Account},\prg{key'}:\prg{Key},\prg{amt}:\prg{nat} }
		{\prg{this}.\prg{\balance}=b-\prg{amt} \wedge \prg{dst}.\prg{\balance}=b'+\prg{amt} } 
		{ \   \prg{this}.\prg{\balance}=b \wedge  \prg{dst}.\prg{\balance}=b'  \  }
}
\\
{\sprepost
		{\strut \ \ \ \ \ \ \ \ \ S_9} 
		{  a:\prg{Account}\wedge
		 \inside{a.\prg{\password}}}
		{\prg{public Account}} {\prg{set}} {\prg{key'}:\prg{Key}}
		{\ \inside{a.\prg{\password} }\ }
		{ \ \inside{a.\prg{\password} } \ }		
}
}
\end{example}
}





 
%??? OR should it be ??? (for the below, we need to add some implications in the Hoare logic for call, or for specs
%\\
%{\sprepost
%		{\strut \ \ \ \ S_7} 
%		{ a:\prg{Account}\wedge  a.\prg{\balance}=b \wedge
%		(\prg{dst}\neq a\neq\prg{this} \vee \prg{key'}\neq a.\prg{\password})}
%	               {\prg{public Account}} {\prg{transfer}} {\prg{dst}:\prg{Account},\prg{key'}:\prg{Key},\prg{amt}:\prg{nat}}
%		{ a.\prg{\balance}=b}
%		{\sdred{ false}}
%}
%\\
%{\sprepostLB
%		{\strut \ \ \ \ \ \ \ \ \ S_8} 
%		{  % b,b':\prg{int} \wedge  
%		\prg{this}\neq \prg{dst}\wedge \prg{this}.\prg{\balance}=b \wedge  \prg{dst}.\prg{\balance}=b' }
%		  {\prg{public Account}}
%		  	  {\prg{transfer}} 
%		   	 {\prg{dst}:\prg{Account},\prg{key'}:\prg{Key},\prg{amt}:\prg{nat} }
%		{\prg{this}.\prg{\balance}=b-\prg{amt} \wedge \prg{dst}.\prg{\balance}=b'+\prg{amt} } 
%		{    false  }
%}
%\\
%{\sprepost
%		{\strut \ \ \ \ \ \ \ \ \ S_9} 
%		{  a:\prg{Account}\wedge
%		 \inside{a.\prg{\password}}}
%		{\prg{public Account}} {\prg{set}} {\prg{key'}:\prg{Key}}
%		{ \inside{a.\prg{\password}}}
%		{ false }
%
%}




\forget{
 \label{example:twostatesarisfy}
\se{We revisit the modules and specifications from Sect. \ref{s:bankSpecEx}, and Example \ref{ex:spacesMore} :}


\begin{tabular}{lllllllll}
$\ModA  \not\models S_1$  &   $\ModA  \models S_2$  &     $\ModA \models S_3$    & $\ModA \models S_5$\\
 $\ModB \not\models S_1$  &   $\ModB \not\models S_2$     &  $\ModB  \not\models S_3$   & $\ModB \not\models S_5$ \\
 $\ModC  \not\models S_1$    & $\ModC \models S_2$ &   & $\ModC \not\models S_3$   & $\ModC \not\models S_5$ 
\end{tabular}
\end{example}
 

 
 \begin{example}
 \label{example:mprepost:sat:one}
 For  %method
 Example \ref{example:mprepostl}, we have
  $\ModA \models S_6$ and $\ModB \models S_6$ and  $\ModC \models S_6$.
Also,  $\ModA \models S_7$ and $\ModB \models S_7$ and  $\ModC \models S_7$.
However,   $\ModA  \models S_8$, while $\ModB  \not\models S_8$.
\end{example}

 \begin{example}
\label{example:mprepost:sat:two}
 For  %method
any   specification  $S \triangleq {\mprepost{A}{p\ C}{m}{x}{C}{A'} }$ and any module  $M$ which does not have a class $C$  with a method $m$ with formal parameter  types ${\overline C}$, we have that $M \models S$.
Namely, if a method were to be called with that signature on a $C$  from $M$, then execution would be stuck, and the requirements from Def. \ref{def:necessity-semantics}(3) would be trivially satisfied.
Thus,   $\ModC \models S_8$. %, even though $\ModC$ does not have a method \prg{set} with the signature given in $S_6$;
\end{example}
}

\subsection{Examples of Semantics of our Specifications}

\begin{example}
 \label{example:mprepost:sat:three}
We  revisit the specifications given in Sect. \ref{s:bankSpecEx},  the three  modules from Sect. \ref{s:bank}, and Example \ref{ex:spesMore}


\begin{tabular}{lllllllll}
$\ModA  \models S_1$  &   $\ModA  \models S_2$ &   $\ModA \models S_3$    & $\ModA \models S_5$\\
 $\ModB \models S_1$  &   $\ModB \not\models S_2$   &  $\ModB  \not\models S_3$   & $\ModB \not\models S_5$ \\
 $\ModC  \models S_1$    & $\ModC \models S_2$ & $\ModC \not\models S_3$   & $\ModC \not\models S_5$ 
\end{tabular}
\end{example}
 

 
 \begin{example}
 \label{example:mprepost:sat:four}
 For  %method
 Example \ref{example:mprepostl}, we have
  $\ModA \models S_7$ and $\ModB \models S_7$ and  $\ModC \models S_7$.
Also,  $\ModA \models S_8$ and $\ModB \models S_8$ and  $\ModC \models S_8$.
However,   $\ModA  \models S_9$, while $\ModB  \not\models S_9$.
\end{example}

 \begin{example}
\label{example:mprepost:sat:five}
 For  %method
any   specification  $S \triangleq {\mprepost{A}{p\ C}{m}{x}{C}{A'} }$ and any module  $M$ which does not have a class $C$  with a method $m$ with formal parameter  types ${\overline C}$, we have that $M \models S$.
Namely, if a method were to be called with that signature on a $C$  from $M$, then execution would be stuck, and the requirements from Def. \ref{def:necessity-semantics}(3) would be trivially satisfied.
Thus,   $\ModC \models S_8$. %, even though $\ModC$ does not have a method \prg{set} with the signature given in $S_6$;
\end{example}

\subsubsection{Free variables in well-formed specifications}
\label{wff:spec:free:more}

We now discuss the requirements about free variables in well-formed specifications as defined in Def. \ref{f:holistic-wff}.
 In scoped invariants, $A$ may only mention   variables introduced by the quantifier, $\overline x$. In method specifications, the precondition, $\overline{x:C'} \wedge A$, may only mention the receiver, \prg{this}, the formal parameters, $\overline y$, and the explicitly introduced  variables, $\overline x$; it may \emph{not}  mention the result  \prg{res}. The postcondition, $A'$, may mention these variables, and in addition, may mention the result, \prg{res}. The mid-condition, $A''$ is about a state which has at least one more frame than the current method's,
and therefore it  may not mention \prg{this}, nor $\overline{y}$, nor \prg{res}.

 \begin{example}[Badly Formed Method Specifications -- Free Variables]
$S_{9,bad}$ is not a well-formed specification, because $a'$ is not a formal parameter, nor free in the precondition. 

 {\sprepost
		{\strut \ \ \ \ \ \ \ \ \ S_{9,bad} }
		{  a:\prg{Account} \wedge  \inside{a} }
		{\prg{public Account}} {\prg{set}} {\prg{key'}:\prg{Key}}
		{   \inside{a}\wedge  \inside{a'.\prg{key}}  }
		{  true }
}
		
% {\sprepost
%		{\strut \ \ \ \ \ \ \ \ \ S_{9,bad\_2} }
%		{  a:\prg{Account} \wedge  \inside{a} }
%		{\prg{public Account}} {\prg{set}} {\prg{key'}:\prg{Key}}
%		{   \inside{a}\wedge  \inside{a'.\prg{key}}  }
%		{  \prg{this}.\balance \ }
%
%}
\end{example}

\subsubsection{Encapsulation in well-formed specifications}
\label{wff:spec:encaps:more}

We now discuss the requirements that post-conditions and invariants as encapsulated -- as defined in Def. \ref{f:holistic-wff}.

 \sdnew{ \begin{example}[Badly Formed Method Specifications -- Postcondition  not encapsulated]
$S_{10,bad}$ is not a well-formed specification,  because the postcondition contains protection in   a negative position ($\neg \inside{\prg{this}.\prg{key}}$), and therefore is  not encapsulated.

 {\sprepost
		{\strut \ \ \ \ \ \ \ \ \ S_{10,bad} }
		{  \ true\  }
		{\prg{public Account}} {\prg{leak}} {\prg{x}:\prg{external}}
		{  \neg \inside{\prg{this}.\prg{key}}   }
		{  true }
}

For similar reasons, $S_{11,bad}$  % and $S_{12,bad}$ are 
is  not well-formed either. The requirement that $\neg \inside{\prg{this}.\prg{key}} $  in the precondition is OK (preconditions need not be encapsulated), but the requirement that $\neg \inside{\prg{this}.\prg{key}} $ in the postcondition is not.

 {\sprepost
		{\strut \ \ \ \ \ \ \ \ \ S_{11,bad} }
		{  \  \neg \inside{\prg{this}.\prg{key}}  \ }
		{\prg{public Account}} {\prg{leak\_preserve}} {\prg{x}:\prg{external}}
		{  \neg \inside{\prg{this}.\prg{key}}   }
		{  true }
}

\vspace{.03cm}

Forbidding specifications like  $S_{10,bad}$ or $S_{11,bad}$ does not restrict the expressiveness of our language. 
They are impossible, or near impossible, to satisfy, because they  talk about properties of the external code, 
over which the internal code has very little  control.

$S_{10,bad}$ is impossible to satisfy: Even if the method \prg{leak} were to pass   to the external object
\prg{x}, there is no way to ensure that \prg{x}, or some other external object will store  \prg{this.key}, and thus no way to guarantee that    $\neg \inside{\prg{this}.\prg{key}}$    will hold  after the call.

On the other hand, $S_{11,bad}$ can  be satisfied, but only if \prg{leak\_preserve} makes no external call.
It \prg{leak\_preserve} were to make an external call, it would be impossible to guarantee that $\neg   \inside{\prg{this}.\prg{key}}$ will be not be invalidated by that external call.
		
\end{example}
}

 \sdnew{ \begin{example}[Badly Formed  Specifications -- Scoped Invariant  not encapsulated]
$S_{12,bad}$ is not a well-formed specification,  because the scoped invariant contains protection in   a negative position ($\neg \inside{\prg{this}.\prg{key}}$), and therefore   is not encapsulated.

$\strut \SPSP\ \ \   S_{12,bad}\ \  \triangleq \ \ \TwoStatesN {\prg{a}:\prg{Account}}  {\neg \inside{\prg{a}.\prg{key}}} $ 

 
Forbidding specifications like  $S_{12,bad}$  does not restrict the expressiveness of our language, because such specifications are impossible to satisfy. Here, $ S_{12,bad}$ requires that external call does not give up access to $\prg{a}.\prg{key}$. Such a guarantee cannot be made by the internal code.
		
\end{example}
}

\input{appendix/appendixExpressivity}
 
