%\se{You use 'some effects' as the opposite of 'uncertain/unknown effects'. I think the effects shouldn't be some effects, but certain effects or specific effects. Some has the feel of randomness (but who knows which) which isn't meant}
%kjx fixed




\section{Introduction}
\label{s:intro}

{\em{External calls}}
%% In today's complex software, internal, trusted, code  is tightly intertwined  with external, untrusted, code: external code
%% calls into internal code, internal code calls
%% out to external code  and external code  even calls back  into internal code --- all within the same call chain.
are pervasive in today's open world software.
% 
% Software is critical to today's open world. 
External, untrusted, or unknown code calls our trusted internal code, that internal code 
calls out to other external code  and external code can even call back  into internal code --- all within the same call chain.
%
%
This paper addresses reasoning about \emph{external calls} --- when
trusted internal code calls out to untrusted, unknown external code.
This reasoning is hard because by ``external code'' we mean untrusted code
where we don't have a specification,
where we may not be able to get source code, or
which may even have been written to attack and subvert the system.
%External code may even have been written by an attacker trying to subvert or destroy the whole system.  
%Worse still, external code can call back into internal code as part of the same call chain. SUSAN:said in the paragraph before.
\forget{
External code may be written years or decades after the core 
trusted system was released --- software plugins for audio and video
editors have lasted for more than thirty years, and new
trusted retrocomputing hosts (``emulators'')
can run code for machines back to the Manchester 
Baby \cite{MAME-SSEM}, where source code is long lost in the bits of
history \cite{Smalltalk-bits}.
 }
 %
%
\vspace*{-2mm}
%
%
\newlength{\curparindent}
\setlength{\curparindent}{\parindent}
\noindent
\begin{flushleft}
\begin{tabular}{@{}lll@{}}
  \begin{minipage}{.55\textwidth}
\setlength{\parindent}{\curparindent}
In the code sketch to the right, 
%  While
an internal module,  $M_{intl}$, has two  methods. %,  \prg{m1}  and \prg{m2}. 
    %contains only trusted code, but
 %   code is trusted,
  Method \prg{m2} takes an untrusted parameter \prg{untrst},
  %and % then 
  at line 6 it  calls an unknown external method \prg{unkn}   passing itself as an argument. 
  The challenge is: % What can that method call do?
   What effects will  that method call have?
What  if  \prg{untrst}   calls back into $M_{intl}$? %, \eg  by calling  \prg{m1}? -- chpped as obvious how. and shorter.
\end{minipage}
& \ \  \   &
\begin{minipage}{.32\textwidth}
\begin{lstlisting}[mathescape=true, language=Chainmail, frame=lines]
module $M_{intl}$        
  method m1 ..
    ...  $\mbox{trusted code}$ ...  
  method m2(untrst:external) 
    ... $\mbox{trusted code}$ ...
    $\mbox{\red{untrst.unkn(this)}}$   
    ... $\mbox{trusted code}$ ...
\end{lstlisting}
\end{minipage}
\end{tabular}
\end{flushleft}

\vspace*{-4mm}
 \paragraph{External calls  need not have  arbitrary   effects} 
If the programming language supports encapsulation (\eg no address forging, private
 fields, %, ownership, no direct memory access 
 \etc) then internal modules can be  written \emph{defensively} %\cite{MillerPhD}, 
so  that effects are either

 \begin{customquote}
\emph{Precluded}, \ie  guaranteed to \emph{never happen}.
  E.g., a  correct  %defensive 
 implementation of the DAO  \cite{Dao} can ensure that  
 the DAO's balance  never falls below the sum of the balances of its subsidiary accounts, or


 \emph{Limited}, \ie  they  \emph{may happen}, but
 only in well-defined circumstances.
%%   \emph{only if} the external 
%% code  uses  certain \emph{object  capabilities}.
E.g., while the DAO does not preclude that a signatory's balance will decrease, it  does ensure that the balance decreases only
as a direct consequence of calls from the signatory.
% if the code causing this reduction uses the signatory itself.
 
 \end{customquote}
 
 \vspace{.1cm}
 
\noindent
%\{Reasoning about  effects limited through capabilities is the remit of this paper.}
Precluded effects are special case of limited effects,
and have been studied extensively in the context of object invariants   \cite{staticsfull,DrossoFrancaMuellerSummers08,BarDelFahLeiSch04,objInvars,MuellerPoetzsch-HeffterLeavens06}.
In this paper, we tackle the more general, %more difficult,  
and more subtle case of reasoning about limited effects for external calls.
%{Reasoning about limited effects
%  %through capabilities
%  is the remit of this paper.}
  
%  While the literature has explored external calls \cite{vanproving,cerise-jacm2024,vmsl-pldi2023,ddd}, ``robust safety'' \cite{gordonJefferyRobustSafety,robustSafetyPatrignani,abate2019journey}, \etc, to our knowledge, there is no widely accepted term limiting the range of effects resulting from external calls.
 


 
 
\paragraph{The Object Capability Model}\footnoteSD{SD says: check whether these two paragraph are the ``necessary'' nature}


The object-capability model combines the capability model
of operating system security \cite{levy:capabilities,CAP}
with pure object-oriented programming
\cite{selfpower,selfexp95,agha_actors_1987}.  Capability-based
operating systems reify resources
%and effects 
%(such as memory regions,
%  files, or even hardware devices)}
as \textit{capabilities} ---
unforgeable, distinct, duplicable, attenuable, communicable bitstrings
which both denote a resource and grant rights over that resource.
%Causing an effect
%on a resource
%requires invoking the capability to that resource:
Effects can only be caused by invoking capabilities:
controlling effects reduces to controlling capabilities.
%% \forget{that reify those
%% effects. To create a new file, for example, a process must possess a
%% capability to the directory that will contain the new file, a
%% capability strong enough to create new files. Invoking the
%% ``\texttt{creat}'' system call \cite{Kern84a} on that capability will
%% create a new file and return a capability for that file.  Duplicating
%% the capability and sending it to another process shares the file with
%% that process, granting that process access to the file; sending a suitably
%% attenuated capability can e.g.\ grant the other process read-only
%% access.}


% \forget{As described by} 
Mark Miller's \cite{MillerPhD}
\textit{object}-capability model treats
object references
as capabilities. 
Building on early object-capability languages such as E
\cite{MillerPhD,ELang} and Joe-E \cite{JoeE}, 
a range of recent programming languages and web systems
\cite{CapJavaHayesAPLAS17,CapNetSocc17Eide,DOCaT14} including Newspeak
\cite{newspeak17},
AmbientTalk \cite{ambientTalk}
Dart \cite{dart15}, Grace \cite{grace,graceClasses},
JavaScript (aka Secure EcmaScript \cite{miller-esop2013}),
and Wyvern \cite{wyverncapabilities} have adopted the object
capability model.
%\forget{Object references cannot be forged; each refers to a distinct object; references can be duplicated (by reference
%assignment, SIMULA's ``\texttt{:-}''); attenuated (typically via Proxies \cite{gof,proxiesECOOP2013}); and communicated to other
%objects. Possession of a ``reference to an object'' (in this setting, also described as possessing an ``object-capability'', a
%``capability'', a ``reference'') can be used to execute any method in that object's public interface. Critically, in object-capability systems
%there is no third-party access controller or security manager or hypervisor which is responsible for security --- rather} 
Security  and encapsulation 
is encoded in the relationships between the objects, and the interactions between them.
%\forget{-- interactions which can communicate capabilities around the system.  As such, object
%capabilities offer a seamless model throughout a programming language or even a distributed global system\footnote{To quote Dan Ingalls,
%``Operating System:\ldots there shouldn't be one'' \cite{ingalls1981}.}. } 
 As argued in  \cite{capeFTfJP}, % by \citet{capeFTfJP}, %Drossopoulou and Noble have put it:
%\textit{``
 object capabilities
 %are a mechanism which makes
 make it possible to write secure
  programs,
% }  %  \cite{capeFTfJP}\forget{ in ways other security models simply do not, }although they 
  but cannot by themselves guarantee that any particular program
  %  using the provided mechanism is, indeed,  secure.
  will be secure.
  %  uses the provided mechanism to be correct or even secure.
%The formulation and verification of such guarantees 
%

%% Object Capability Model (OCAP)  \cite{MillerPhD}  is a security model   based on the concept of capabilities, which  
%% Capabilities are transferable rights that allow the performance of one or more operations on a specific object. 
%% Capabilities are unforgeable and can be obtained only through creation of the corresponding object, or through receipt of a message.
%% Capabilities are {\emph{necessary}} conditions for causing effects; callers can only produce effects if they possess the required capabilities. 
%% For example, a signatory can withdraw funds from a DAO only if they hold a ``withdraw'' capability for that specific account within that particular DAO.  

%% Languages like E, subsets of JavaScript (CITE SES), Caja, Wyvern, Rholang,  implement OCAP principles. CITE-CITE-CITE


\paragraph{Reasoning with Capabilities}
 Recent work has developed logics to prove properties of programs employing object capabilities.
\citet{ddd} develop a logic to prove that code employing object
capabilities for encapsulation preserves invariants for
intertwingled code, but without external calls. 
%\forget{Relying on techniques from separation logic, they intend their proofs e.g.\ to support library components that can be reused in different contexts.}
%\forget{--- which is implicitly support in our distinction between external and internal code.
% \footnote{\red{TO CHECK}\kjx{I have looked at it - good enough for
% me.}}}
\citet{dd}
%\forget{work with Kripke logical relations describing program semantics, which can}
describe and verify invariants
about multi-object structures and the availability and exercise of object capabilities.  %\citeauthor{dd} 
% They propose and solve several
% problems involving external calls. 
%\forget{,  but require proofs specific to each individual problem, rather than a systematic Hoare logic.}
Similarly,
\citet{vmsl-pldi2023} propose a separation logic to support formal modular reasoning about communicating VMs, including in the presence of unknown VMs.
\citet{irisWasm23} specify WASM modules, %and compose them in a simple target language, 
and prove that adversarial code can affect other modules only through functions they explicitly export. 
%Since external calls may in turn call any number of internal, public methods, 
%it is crucial  to prove that \taming of effects is preserved by the module's 
%\emph{emergent} behaviour, \ie any intertwining of external code with calls to internal, public methods.
\citet{CassezFQ24} model external calls as an unbounded number of invocations to a module's public interface.
 
The approaches above %, with partical exception of  \cite{{FASE,OOPSLA22} 
do not aim to support
%indirect, eventual access to capabilities.
general reasoning about external effects limited through capabilities.
% to limit the effects of external calls to untrusted code.
\citet{FASE} and \citet{OOPSLA22} begin to tackle external effects; %concentrate on access to capabilities;
the former proposes  ``holistic specifications'' to describe a module's emergent behaviour. 
and the latter develops 
a tailor-made logic to prove that modules which do not contain external calls adhere to holistic specifications.
%They propose a tailor-made proof for an external call.
%We argue that in order to \emph{prove} that effects are \tamed,
%we need a way to specify the conditions necessary for that effect, 
%and a way to prove that the   module in its entirety preserves this \taming.
%In order to prove that the module in its entirety preserves this \taming, 
% \vspace{.2cm} 
% Thus, no work so far has tackled Hoare logics which support external calls as well as \tamed effects.
%
%\vspace{.03cm}
%
%Other than this, 
%{To} reason about the effects of ific external calls,
%the systems above {would} rely on problem-specific, custom-made proofs rather than a Hoare logic.
Rather than relying on problem-specific, custom-made proofs, we propose a Hoare logic that addresses access to capabilities, limited effects, and external calls.
% and their limited effects.

%\paragraph{This paper's contributions}
\paragraph{This paper contributes}
% (0) a demonstration that object-capabilities can reify the effects of external calls,
(1) \emph{protection assertions} to limit access to object-capabilities, % and concomitantly limit effects}, 
(2) a specification language to define how limited access to capabilities should limit effects, %effects should be limited by capabilities.
(3) a Hoare logic to reason about external calls and to prove that modules satisfy their %\tamed effects 
specifications,
(4) proof of soundness,
(5) a worked illustrative example {with a mechanised proof in Coq}.
 

  \paragraph{Structure of this paper}
Sect.\ \ref{s:outline}   outlines the main ingredients of our approach in terms of an {example}.
% chopped to not get a page break:, and explains the main contributions.
Sect.\ \ref{sect:underlying} outlines a simple object-oriented language used for our work. 
% Sect. \ref{s:auxiliary}  {contains essential concepts for our study}.
Sect.\ \ref{s:assertions} and 
 Sect \ref{sect:spec}  give syntax and semantics of assertions  and  specifications.
%{while Sect. \ref{s:preserve} discusses preservation of satisfaction of assertions.}
Sect.\ \ref{sect:proofSystem} develops a Hoare logic %triples and quadruples
 to prove external calls,  that a module adheres to its %\tamed effects
 specification, and summarises  the Coq proof
  of  our running example (the source code will be submitted as an artefact). 
Sect.\ \ref{sect:sound:proofSystem} outlines {our} proof of soundness of
the Hoare logic. 
  %Sect.\ \ref{sect:related} discusses related
%work and
 Sect.\ \ref{sect:conclusion} concludes with related work. 
Fuller technical details can be found in the appendices in the accompanying materials. 
