        
This is a very scrappy first draft of a related work section ---
hopefully someone will at least edit it, fix typos, and remove this
sentence! It is Sophia's text from the introduction, plus stuff from
previous papers (notably the TOSEM rejected draft, including verona,
necesity, holistic) and various other grant applicatiosn whot I wrote.


\paragraph{Behavioural Specification Languages} 

\citet{behavSurvey2012} provide an excellent survey of
contemporary specification approaches.  With a lineage back to Hoare
logic \cite{Hoare69}, Meyer's Design by Contract \cite{Meyer97} was the
first popular attempt to bring verification techniques to
object-oriented programs as a ``whole cloth'' language design in
Eiffel.  Several more recent specification languages are now making
their way into practical and educational use, including JML
\cite{Leavens-etal07}, Spec$\sharp$ \cite{BarLeiSch05}, Dafny
\cite{dafny} and Whiley \cite{whiley15}.
. Our approach builds upon
these fundamentals, particularly Leino \& Shulte's
%\kjx{and Naumann's} 
formulation of
two-state invariants \cite{usingHistory}, and Summers and
Drossopoulou's Considerate Reasoning \cite{considerate}.
%
In general, these approaches are within the setting of a closed
system, where modules can be trusted to co{\"o}perate. In this paper
we aim to
% illustrate the kinds of techniques required
work
in an open system where many modules are unknown by, or even actively
antagonistic to the system as a whole: and even worse, where
known / trusted / internal code must interact with unknown / untrusted
/ external code, both making outgoing calls into the unknown components,
and recieving incoming calls from the unknown components --- even
callbacks, i.e.\ incoming calls back into internal components while
those internal components executions are themselves suspending during
the outgoing calls\ldots

Since open calls may in turn call any number of internal, public
methods, it is crucial to prove that \taming of effects is preserved
by the module's \emph{emergent} behaviour, \ie any series of incoming
external calls, intertwined with internal calls to internal methods,
and then outgoing external calls either to other external objects (if
we're lucky) or callbacks into the internal objects that are currently
suspended waiting for the results of the external call (if we're not).

\citet{CassezFQ24} illustrates one approach to handling external
calls in the context of the Dafny verifier. The key idea is that
external callbacks (outgoing external calls that themselves make
re{\"e}ntrant incoming calls) are necessarily restricted to the public
interface of the module to be verified.  This technique replaces
actual external calls with a generated \texttt{externalcall()} method,
where that method nondterministically issues a squence of calls to the
module's public interface.  This means that in order to verify their
code, programmers must ensure that no such sequence of external calls
will break their invariants, and provide enough information for Dafny
to accept that proof.

%% \citet{Rinard03} first made clear how type systems express
%% properties that hold for all program executions: type safety assures
%% that e.g.\ typesafe execution can continue indefinitely, even if the
%% data being processed, memory layouts, or communications into and out
%% of a program are completely randomised. The point of program
%% verification, of course, is to give rather more semantic guarantees.



%\footnote{Properties which rely on the type system, are of course preserved,%%  but what about more semantic properties?}
%% \textbf{xxx-OOPSLA-22-workshop} SOPHIA WHAT IS THIS REFERENCE? does this by
%% replacing an open call though an unbounded number of calls to a
%% module's public method.

\citet{FASE} instead proposes holistic specifications to describe a
module's emergent behaviour. Holistic specifications take an external
perspective across methods in one or more modules' interfaces
(compared with pre- and post- conditions that supply internal
specifications for individual method implementations, and class,
monitor, or module invariants that make intenal specifications for all
the methods in a single class, monitor, or module.
\citet{OOPSLA22} builds upon this work, offering a specification
language based on necessary conditions, rather than sufficient
conditions as in most specification languages, and a Hoare logic to 
that can prove modules  adhere to such specifications: unfortunately
neither of these systems support any kind of external calls --- rather
they rely on simplistic type-based module confinement
\cite{ownership-confinement-jacm2004,confined-toplas2007} which
prevents any kind of external references or external calls, either
outgoing or incoming.

Key to these approachces is the goal of proving properties that must
hold across every execution. The semantics we use in this paper,
defined in section~\ref{def:necessity-semantics}, are based on
primitive classical assertions that characterise individual states.
The semantics of the underlying language require those
specifications to hold over all states that can possible arise from
any execution of a module beginning from a defined initial state

Hyper-Hoare Logic \cite{hyper-hoare-pldi2024} embodies 
an alternative formulation of the problem. Hyper-Hoare logic expresses
programs' hyper-properties in terms of hyper-assertions directly over
sets of states, rather than e.g.\ our classical approach. This
increases the power of the specification language
although at the cost of making the specification languaghe more complex.

\paragraph{Guarantee Reasoning}

Rely-Guarantee
\cite{relyGuarantee-HayesJones-setss2017,relyGuarantee-vanStaden-mpc2015}
and Deny-Guarantee \cite{DenyGuarantee} reasoning techniques
distinguish between assertions guaranteed by a thread, and those a
thread can reply upon (rely-guarantee) or actions denied to all other
threads (deny-guarantee, where the ``deny'' condition limits access to
some state to a particular thread, i.e.\ a thread can rely that other
threads will not write to the denied state).
}
Assume-guarantee
reasoning, especially circular
assume-guarantee  \cite{circular-assume-guarantee-fm2015} evinces
another commonality with our work here: one of the reasons callbacks
are hard to handle in general is that circular control flow can lead
all too easily towards circular reasoning.
%%%
The Hoare quadruples that we use in our work are clearly related to
the ``guarantee'' portion of rely- and deny- guarantee. When a
specification includes a guarantee, that guarantee must be maintained
by every ``atomic step'' in an execution
\cite{relyGuarantee-HayesJones-setss2017}, rather than just at method
boundaries as in visible states semantics
\cite{MuellerPoetzsch-HeffterLeavens06,DrossoFrancaMuellerSummers08,considerate}.
Liquid information flow \cite{liquid-infoflow-icfp2020} and liquid
resource types \cite{} are likewise similar to the work in this paper.

In concurrent reasoning, the original domain of Rely- and
Deny-Guarantee reasoning, this is because shared state may be accessed
by another co{\"o}perating thread at any time, not just outside method
execution boundaries: while in our case, it is because unprotected
state may be accessed by an untrusted component within the same
thread.  The guarantees correspond to our requirements that certain
properties be preserved by all code linked to the current
module. Compared with our work, deny-guarantee assumes
co{\"o}peration: composition is legal only if threads adhere to their
guarantees. In this work, we use these ``guarantees'' precisely to
ensure our code can interoperate safely with external untrusted code
irrespective of whatever the untrusted code does.






%% %Finally, 
%% \sd{Our} work is also related to the causal obligations in Helm et
%% al.'s behavioural contracts \cite{helm90}. Causal obligations allow
%% programmers to specify e.g.\ that whenever one object receives a
%% message (such as a subject in the Observer pattern having its value
%% changed) that object must send particular messages off to other objects
%% (e.g.\ the subject must notify its observers). \Chainmail's control
%% %SD: not "control flow"
%%  operator % (`$\Calls{\_} {\_} {\_} {\_} $) 
%%  %allows  programmers to make
%%  \sd{supports}  similar specifications, (e.g. 
%%  ${\Calls{\_}  {\prg{setValue}} {\prg{s}} {\prg{v}}}  \rightarrow \Future{\Calls{\prg{s}}{\prg{notify}}{\prg{s.observer}}{\prg{v}}}$ --- when a subject receives a \prg{setValue} method,
%%   it must ``forward'' those messages to the observer.

\paragraph{Defensive Consistency}

%cute but wrong.
%To misparaphrase Tolstoy, secure systems are all alike;
%every insecure system is insecure in its own way
%\cite{WikipediaAnnaKareninaPrinciple}.

In an open world, we cannot rely on the kindness of strangers: rather
we have to ensure our code is correct regardless of whether it
interacts with friends or foes.
Attackers 
\textit{``only have to be lucky once''} while secure systems 
\textit{``have to be lucky always''} \cite{IRAThatcher}.
% SD 
Miller \cite{miller-esop2013,MillerPhD} defines the necessary approach
as \textbf{defensive consistency}: \textit{``An object is defensively
  consistent when it can defend its own invariants and provide correct
  service to its well behaved clients, despite arbitrary or malicious
  misbehaviour by its other clients.''}  Defensively consistent
modules are particularly hard to design, to write, to understand, and
to verify: but
% they have the great advantage that
they make it much
easier to make guarantees about systems composed of multiple components
\cite{Murray10dphil}.

\textit{kjx: could talk about Robust Safety and other terms
  otherpeople use here?}



\paragraph{Object Capabilities and Sandboxes.}
{{\em Capabilities} as a means to support the development of concurrent and distributed system  were developed in the 60's
by Dennis and Van Horn \cite{Dennis66}, and were adapted to the
programming languages setting in the 70's \cite{JamesMorris}. 
{\em Object capabilities} were first introduced~\cite{MillerPhD} in the early 2000s},
 and many recent % work attempts to manage
studies manage
to verify  safety or correctness of object capability programs.
Google's Caja \cite{Caja} applies   sandboxes, proxies, and wrappers
 to limit components'
access to \textit{ambient} authority.
% --- that is, capabilities that
%can be obtained from the wider environment, rather than being granted
%to a component explicitly.
Sandboxing has been validated
formally: Maffeis et al.\ \cite{mmt-oakland10} develop a model of
JavaScript, demonstrate that it obeys two principles of
object capability systems
%  (``connectivity begets connectivity'' and
%``no authority amplification''), and then % uses these principles to
and show  how untrusted applications can be prevented from interfering with
the rest of the system. 
Recent programming languages % and web systems
\cite{CapJavaHayesAPLAS17,CapNetSocc17Eide,DOCaT14} including Newspeak
\cite{newspeak17}, Dart \cite{dart15}, Grace \cite{grace,graceClasses}
and Wyvern \cite{wyverncapabilities} have adopted the object
capability model.

%% \paragraph{Verification of Dynamic Languages}
%% A few formal verification frameworks  address JavaScript's highly
%% dynamic, prototype-based semantics. Gardner et al.\ \cite{Gardner12}
%%  developed a formalisation of JavaScript based on separation logic
%% % that they have used
%% and verified   examples. Xiong and Qin et
%% al.\ \cite{XiongPhd,Qin11}  worked on similar lines.
%% % More substantially,
%% Swamy et al.\ \cite{JSDijkstraMonad}  recently
%% developed a mechanised verification technique for JavaScript based on
%% the Dijkstra Monad in the F* programming language.  Finally, Jang et
%% al.\ \cite{Quark} % have %  managed to provide
%% developed a machine-checked proof of
%% five important properties of a web browser --- again similar to our
%% % \prg{any\_code} 
%% invariants --- such as
%% % \textit{``no tab may interfere with
%% %  another tab''} and 
%% \textit{``cookies may not be shared across
%%   domains''} by writing the minimal kernel of the browser in Haskell.
  
%%   \paragraph{JavaScript analyses.}
%% More practically, 
%% Karim et al. apply static analysis on
%% Mozilla's JavaScript Jetpack extension framework \cite{adsafe}, including
%%  pointer analyses. % In a different direction,
%% Bhargavan et al.\ \cite{DefJS}
%% extend language-based sandboxing techniques to support defensive
%% components that can execute successfully  in otherwise untrusted
%% environments.   Politz et
%% al.\ \cite{ADsafety} use a JavaScript type checker to check
%% properties such as
%% % \textit{``widgets cannot obtain direct references
%%  % to DOM nodes''} and
%%  \textit{``multiple widgets on the same page
%%   cannot communicate.''}
%% % --- somewhat similar in spirit to our \textbf{Pol\_4}.
%% Lerner et al.\ extend this system to ensure browser
%% extensions observe \textit{``private mode''} browsing conventions,
%% such as that \textit{``no private browsing history retained''}
%% \cite{Lerner2013b}.  Dimoulas et al.\ \cite{DPCC14} generalise the
%% language and type checker based approach to enforce explicit policies,
%% % although the policies  are restricted to
%% that  describe  which components  may
%% access, or may influence the use of, particular capabilities.
%% Alternatively, Taly et al.\ \cite{secureJS}
%% model  JavaScript APIs in Datalog, and then
%% carry out a Datalog search for an ``attacker'' from the set of all
%% valid API calls. 



\paragraph{Verification of Object Capability Programs}

Murray made the first attempt to formalise defensive consistency and
correctness~\cite{Murray10dphil}.  Murray's model was rooted in
counterfactual causation~\cite{Lewis_73}: an object is defensively
consistent when the addition of untrustworthy clients cannot cause
well-behaved clients to be given incorrect service.  Murray formalised
defensive consistency very abstractly, over models of (concurrent)
object-capability systems in the process algebra CSP~\cite{Hoare:CSP},
without a specification language for describing effects, such as what
it means for an object to provide incorrect service.  Both Miller and
Murray's definitions are intensional, describing what it means for an
object to be defensively consistent.


\citet{capeFTfJP,capeFTfJP14} have
analysed \citet{MillerPhD}'s Mint and Purse example 
% SD Chope details by
% expressing it in Joe-E 
% a Java subset without reflection and static
%fields, 
%and in Grace \cite{capeFTfJP14}, 
and discussed the six
capability policies 
% that characterise the correct behaviour of the
% program, 
as proposed by \citet{MillerPhD}.
%We argued that these policies require a novel
%approach to specification, and showed some first ideas on how to use
%temporal logic.
%  an unpublished technical report
\citet{WAS-OOPSLA14-TR} %, {they} % Drossopoulou and Noble
sketched a  specification language,  \sd{used}  it to  
specify the six policies from \cite{MillerPhD}, % however,
%{their} partial formalisation showed that % they allowed
\sd{showed} that several possible interpretations were possible, %.  They also 
\sd{and} uncovered
the need for another four further policies.
%  and formalised them as well, showing how different implementations of the underlying Mint and Purse
% systems coexist with different policies \cite{capeIFM14},
They also
  sketched how 
a trust-sensitive 
example (the escrow exchange) could be verified in an open world
\cite{swapsies}. 
% In contrast, our work focuses on the semantics of the  \Chainmail\ specification
% language and how it can be used to provide holistic specifications for
% robust programs.
%% \sd{Their work does not support the concepts of control, time, or space, as in \Chainmail,
%%   but it offers a primitive expressing trust.}

 
Some more recent work is particularly relevant to proof of properties
of programs employing object capabilities in this open world.

% SWASSEY 18 garg dreyer

\citet{ddd}  designed OCPL, a logic
for object capability patterns, that supports specifications and
proofs for object-oriented systems in an open world.  
% The key idea here is to 
\sd{They} % say it simpler
draw on verification techniques for security and
information flow: separating internal implementations (``high values''
which must not be exposed to attacking code) from interface objects
(``low values'' which may be exposed).  OCPL supports defensive
consistency % (Swasey et al.\ use 
(\sd{they} use the term ``robust safety'' from the
security community \cite{Bengtson}) via a proof system that ensures
low values can never leak high values to external attackers. 
%\susan{How does this imply that high values can be exposed?}
%\james{typo fixed: it's LOW values that can be exposed}
This means that low values \textit{can} be exposed to external code,
and the behaviour of the system is described by considering attacks only
on low values.  %OCPL is a program logic, and Swasey
\sd{They} use that logic to
prove a number of object-capability patterns, including
sealer/unsealer pairs, the caretaker, and a general membrane.

This work was then developed to prove the memory properties of Rust in
the RustBelt system \cite{RustBelt18}, based around the IRIS automated
separation logic \cite{iris-jfp2018}.  RustBelt was combined with the
RustHorn verification system that uses classical logic 
\se{SUSAN: doesn't it use Horn Clause Logic? or Constraint Horn Clause Logic?} to verify
programs in the safe subset of Rust \cite{RustHorn-toplas2021},
producing RustHornBelt \cite{RustHornBelt-pldi2022} that can verify
both safe and unsafe Rust programs in classical first-order logic.
Similar techniques were extended to C \cite{RefinedC-pldi2021}:
note that while C is not ``memory-safe'', neither is unsafe Rust.
Compared to the work we present in this paper, both combine ``safe''
(internal) and ``unsafe'' (external) code: however both projects have
rather different interpretations of ``safety''.  These Rust projects'
``unsafe'' code is code that is not known to the verification system,
is trusted, and indeed will be verified, but for which the language
does not enforce its characteristic memory constraints, whereas in our
work here, both internal and external code is ''memory safe'' (garbage
collected, object capabiltiies cannot be forged) but the unsafe code
is unknown to the verifier, and may be directly antagonistic to the
overall properties of the system.



\citet{schaeferCbC} have recently
% taken a similar approach to Swasey,
% adding support for
\sd{added}  support for information-flow security % in a setting 
\sd{using} refinement to ensure correctness (in this case confidentiality) by
construction. 
% Although designed to support
% confidentialty, it seems likely that the information-flow guarantees
% could also be used to ensure robustness.  





% devrise birkedal S&P 2016; 
 
\citet{dd}  have deployed
   \sd{powerful} %rather more complex
  theoretical techniques to address similar problems:  % Devrise et al.\ 
  \sd{They} show how step-indexing, Kripke worlds, and representing objects
as state machines with public and private transitions can be used to
reason about % object-oriented programs in general.
\sd{object capabilities}.
Devriese have demonstrated solutions to a range of exemplar problems,
including the DOM wrapper and a mashup application.
% Although the formal techniques are much more sophisticated than we
%apply here, and consequently 
% not true can e.g.\ reason about recursion where we
%cannot, there are some similarities, e.g.\ with the 
\sd{Their} distinction
between public and private transitions %being related 
\sd{is similar} to the
distinction between internal and external objects.


This stream of work has culminated in VMSL, a seperation logic for
virtual machines 
\cite{vmsl-pldi2023}
(sadly not a logic for VAX/VMS systems)
that demonstrates how VMs can assume ``robust safety'' (more or less
defensive consistency), and Cerise, a full-scale extension build on 
seperation logic based on capability-safe CPUs \cite{cerise-jacm2024}.
(Both VMSL and Cerise are also build with Iris.)
Cerise uses Iris invariants to support proofs of programs with both
outgoing external calls and even incoming callbacks,
however this requires a relatively low-level proof in the Iris variant
of seperation logic. 



%% Devriese - 17? also develop xx logic for object capabilties and code invariants, they propose interesting problems for open calls, 
%% but can only prove these properties through tailor-made proofs rather than Hoare logics.



%% Dreyer OOPSLA'18 develop a logic to prove that   code employing object capabilities as encapsulation features preserve invariants for intertwined code, but in the absence of open calls. 
%%  \footnote{\red{TO CHECK}}.


 
The work we present here differs from Swasey, Schaefer's, and Devriese's work in a number of ways:
% \citet{ddd} and \citet{schaeferCbC} 
\sd{They} are primarily concerned \sd{with} %about
mechanisms that ensure encapsulation (aka 
confinement)
while
in this work we focus more on specifications of programs' behaviour,
and constructing proofs that programs can meet those specifiations.
%% \Chainmail\ abstracts away from any mechanism via
%% an external predicate. 
\sd{Swasey and Schaefer use powerful mathematical techniques, such as
  Kripke worlds and step-indexing  
which  the users need  to understand in order to write their specifications,
while \Chainmail users only need  to understand  first order logic and 
their additional holistic operators.}
% While \Chainmail's $\Using{}{}$ is related to Banerjee
% and Naumann's region sets, the assertion languages here are mostly
% traditional (extensions of) Hoare logics --- Swasey et al.\ build on a
%concurrent separation logic. 
%% \sd{ Finally, none of these systems offer the kinds of
%% holistic assertions addressing control flow, change, or temporal
%% operations that are at the core of \Chainmail's approach.
%% }

Scilla \cite{scillaOOPSLA19} is a minimalistic typed functional

bytecode. Scilla's semantic model is restricted, assuming actor based
communication and restricting recursion,  thus facilitating static
analysis of Scilla contracts and ensuring termination.
Scilla is able to demonstrate that a number of popular Ethereum
contracts avoid type errors, out-of-gas resource failures, and
preservation of virtual currency. 
Scilla's semantics are defined formally, but have not yet been represented in a
mechanised model.

%% \kjx{NPChecker \cite{NPcheckerOOPSLA19} analyses Ethereum smart
%% contracts to detect bugs related to nondeterministic
%% execution. NPChecker undertakes an information flow
%% analysis to detect potential read-write hazards
%% particularly reentrancy and ordering dependencies.
%% \textbf{We don't do concurrency. Do we need this one? I don't think so}
%% }



\textbf{SAY SIOMETHNG
ABOUT  \citet{rich-specs-smart-contracts-oopsla2021}}


Finally, the recent VerX tool is able to verify a range of
specifications for solidity contracts automatically \cite{VerX}.
VerX's specification language is based on
temporal logic, 
but only within a past modality, while e.g.\ \citet{OOPSLA22} has both past
and future modalities.
%% VerX specifications can also include
%% predicates that 
%% model the current invocation on a contract (similar to \Chainmail's
%% ``calls''), can access variables, and compute sums (only) over
%% collections.
\Chainmail\ is strictly more expressive as a
specification language,
%% including quantification over objects and sets
%% (so can compute arbitrary reductions on collections) and of course
%% specifications for permission (``access''), space (``in'') and
%% viewpoint (``external'')
%% which have no analogues in VerX. 
%% Unlike
%
however unlike
%
our work, VerX includes a practical tool that has
been used to verify   a hundred properties across case studies of
twelve Solidity contracts.
%\textbf{(ideally also say something about proof status)}}
 

Unlike systems like Cerise, which can handle callbacks, VerX
is restricted to ``effectively call-back free'' programs
\cite{Grossman,Albert,relaxed-callbacks-ToDES}, which assume any callbacks can be delayed until after the
original incoming call to the internal object has finished 
--- thus turning an incoming \textit{callback} (a call invoked 
to handle an outoing external call) into a more straightforward plain
incoming call. 


The \textsc{ConSol} project \cite{consolidating-pldi2024} takes a
different approach, drawing on programming langauge contracts that are
checked at runtime \cite{FinFel01}. \textsc{ConSol} provides a
specification langauge, based around classical pre- and post-conditions and
a special higher-order treatment of the blockchain addresses used in
smart contracts. The project analyses twenty well-known erroneous
contracts and claims that ConSol specifications would prevent those
errors.
SCIO* \cite{secure-io-fstar-popl2024} implemented in the dependently
typed langauge F* offers another approach: the system supports both
verified and unverified code --- with unverified code 
corresponding to our external or untrusted modules.  SCIO,
like \textsc{ConSol}, then uses runtime checks in the form of
higher-order contracts and reference monitors to ensure the unverified
code cannot break invariants on the verified code. Both these
approaches are similar to e.g.\  techniques for gradual
verification \cite{gradual-verification-popl2024,Cok2022} which often
rely on inserting dynamic checks at the interface between verified
(known, trusted) and unverified (unknown, untrusted) code.



\paragraph{Incorrectness Logic.} O'Hearn~\cite{IncorrectnessLogic} defined a Hoare
Logic for modelling program incorrectness. O'Hearn's Incorrectness Logic
is based on a Reverse Hoare Logic \cite{reverseHoare}, and empowers programmers to 
specify preconditions under which specific errors and program states may result. 
Incorrectness Logic provides a sound and compositional way to reason about 
the presence of bugs rather than the absence of bugs. 
As with Hoare logic, Incorrectness Logic provides a system
for reasoning about sufficent conditions for post-conditions to hold.
However, where Hoare logic specifies the shape of the result of execution 
of all program states that satisfy the precondition, Incorrectness Logic
specifies that all states that satisfy the postcondition are reachable
from those that satisfy the precondition. This suits the specification
of program errors, as it allows for the exclusion of false negatives.


%% In comparison, \Chainmail, as with Hoare Logic, is concerned with correctness
%% (as seen in the exemplars of Section \ref{sect:problemdriven}). 
%% Extending the comparison, \Chainmail differs from both Hoare Logic and Incorrectness,
%% in the ability to specify, not just sufficient conditions, but necessary conditions for 
%% reaching certain program states. Neither Incorrectness Logic, 
%% nor Hoare Logic allows for such specifications.


In comparison, this work, as with traditional Hoare Logic, is
concerned with correctness rather than incorrectness.
% (as seen in the exemplars of Section \ref{sect:problemdriven}). 
Extending the comparison,
this work 
differs from most Hoare Logics and Incorrectness Logics,
which require specifications of entire closed systems,
with the ability to specify open systemss that support incoming calls
out of untrusted components, outgoing calls back into untrusted
components, and even callbacks where
our verified modules call out to unknown and untrusted components,
which then call back into our verified and trusted components.


% SD chopped as did not like
%As with Swasey et al.\ this work does not provide a holistic
%assertion language like \Chainmail.
% SD Chopped, as it sounds as if their is not real code, which is debatable
% and what is an extensional framework? they would say that theirs is too.
%In contrast, \Chainmail\ is
%meant for describing and reasoning about real code, and we provide an
%expressive, extensional framework for evaluating defensive consistency
%in actual open systems.
%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%NOTES:
%% the other thing this section needs to do, particularly with Devrise, is to lay out precisely the way our work is more limited than theirs:
%% (Swasey, I'm more and more convinced, is just ownership-via-a-proof-system) 
%% we don't step-index, don't have logical relations, etc: what do we lose by NOT having those things
%% (or what do we gain by having those things...

%% The "deep" comparison with Swasey and with Devirese (and also
%% information flow control and temporal logics) needs to say why these
%% works are not as good (expressive? easy to understand?) as ours.
%% Currently the Related work just mentions them, but does not answer the
%% question as to why our work is important when theirs already has been
%% published.




%% *Difference between Spec Languages and Chainmail*  One way to tackle
%%  this would be to enumerate which elements of Chainmail appear at
%%  other works, which do not, and claim that Chainmail’s novelty is the
%%  good combination of these elements


%% Eg: reflection about contents of stack and heap (in classical Hoare
%% Logics), two state assertions (JML etc), invariants (Hoare and Meyer),
%% internal/external (Liskov?, Noble et al,modules in Neumann and also
%% O’Hearn), time (temporal logic, but they do not have the other stuff),
%% Control (none?), Space (in Sep. logic, and in effects, buyt the
%% meaning is different), Permissions (our earlier work, and less
%% flexible approaches such as owenrship types and perhaps also
%% oinformation flow control), Authority (effect systems and modifies
%% clauses, and perhaps also Bierman&Parkison abstract predicates, but
%% there it is tied to pre-post conditions.


%% Also, point out difference between our invariants and Hoare
%% triples. Subtle and needs thinking







%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Neither effort addresses specification languages for security and
%% robustness, provides Hoare logics to reason about object-capability
%% programs.

%% , model protocols that dynamically ascribe trust
%% \cite{swapsies,lefthand} or quantify the damage an untrustworthy
%% object can do.






% \kjx{History-Based Specification and Verification of Scalable
%  Concurrent and Distributed Systems --- ICFEM15}


% \paragraph{Specifying Design Patterns}
% Techniques for specifying Design Patterns go back at least to 
% Helm's contracts \cite{Helm92}.

% more importantly: work on formalisation of design patterns.
% (again look at JC grant, even if refs are 5 years old)
% let's be shameless here...



% This search is similar to the quantification over
% potential code snippets in our model.
% The problem posed by the Escrow example is that it establishes a two-way
% dependency between trusted and untrusted systems --- precisely the
% kind of dependencies these techniques prevent.

% %These approaches are all based on static analyses.
%  The WebSand
% \cite{flowcaps11,sabelfeld-inlining2012} and Jeeves \cite{jeeves2012}
% projects use dynamic techniques to monitor safe execution of information flow policies.
%  Richards et al.\ \cite{FlacJS}   extended this approach by
% incorporating explicit dynamic ownership of objects (and thus of
% capabilities) and policies that may examine the history of objects'
% computations. While these dynamic techniques can restrict or terminate
% the execution of a component that breaches its security policies, they
% cannot guarantee in advance that such violations can never happen.
% While information flow policies are concerned with the flow of objects (and thus also capabilities)
% across the program code, our work is more concerned with the identification of the objects which protect
% the services.

%Compared with all these approaches, our work   focuses on
%\textit{general} techniques for specifying (and ultimately verifying)
%capability policies, whereas these systems are generally much more
%\textit{specific}: focusing on one (or a small number) of actual
%policies. % This seems to be because contemporary object capability
%programming is primarily carried out in JavaScript, but
% There are few

 
% \paragraph{Relational models of trust.}
% Artz and Gil \cite{artz-trust-survey-2007} survey various
% types of trust in computer science generally, although trust has also
% been studied in specific settings, 
% %
% ranging from peer-to-peer systems \cite{aberer-trust-p2p-2001} and
% cloud computing \cite{habib-trust-cloud-2011} 
% to mobile ad-hoc networks \cite{cho-trust-survey-adhocnets-2011}, 
% the internet of things \cite{lize-trust-IoT-2014}, 
% online dating \cite{norcie-trust-online-dating},
% and as a component of a wider socio-technical system
% \cite{cho-trust-sustainable-2013,walter-trust-cloud-ecis2013}. 
% %
% Considering trust (and risk) in systems design, Cahill et al.'s overview
% of the \textsc{Secure} project \cite{cahill-trust-pervasive-2003}
% gives a good introduction to both theoretical and practical issues of
% risk and trust, including a qualitative analysis of an e-purse
% example. This project builds on Carbone's trust model
% \cite{carbone-formal-trust-2003} which offers a core semantic model of
% trust based on intervals to capture both trust and uncertainty in that
% trust. Earlier Abdul-Rahman proposed using separate relations for
% trust and recommendation in distributed systems
% \cite{abdul-rahman-distributed-trust-1998}, more recently Huang and
% Nicol preset a first-order formalisation that makes the same
% distinction \cite{huang-formal-semantics-trust-calculus-2010}.
% Solhaug and St{\o}len \cite{solhaug-trust-uncertainty-2011} 
% consider how risk and trust are related to uncertainties over
% actual outcomes versus knowledge of outcomes.
% Compared with our work, these approaches produce models of trust
% relationships between high-level system components 
% (typically treating risk as uncertainty in trust) 
% but do not link those relations to the system's code. 



% \paragraph{Logical models of trust.}
% \sd{A detailed study of how web-users decide whether to trust appears in \cite{GilArtz}.}
% \sd{Starting with \cite{Lampson92},} various different logics have been used to measure trust in different
% kinds of systems.
% Murray and Lowe \cite{murray10-infoflow} model object capability
% programs in CSP, and use a model checker to ensure program executions
% do not leak authority.
% Carbone et al.\ \cite{carbone-formal-trust-2011}
% use linear temporal logic to model specific trust relationships in service
% oriented architectures.
% Ries et al.\ \cite{habib-trust-CertainLogic-2011} evaluate trust under
% uncertainty by evaluating Boolean expressions in terms of real values
% for average rating, certainty, and initial expectation.
% % Perhaps closer to our work, Aldini
% Aldini \cite{aldini-calculus-trust-IFIPTM2014} describes a temporal logic for
% trust that supports model checking to verify some trust properties.
% Primiero and Taddeo \cite{primiero-modal-theory-trust-2011} have
% developed a modal type theory that treats trust as a second-order
% relation over base relations between
% counterparties. Merro and Sibilio
% \cite{merro-calculus-trust-adhoc-facs2011} developed a trust model for
% a process calculus based on labelled transition systems.
% Compared with our proposal, these approaches use
% process calculi or other abstract logical models of systems, rather
% than engaging directly with the system's code.






%%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% 
%%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% %%%% 







\paragraph{Protection vs.\ Ownership}

Finally it is interesting to compare our approach to tamed effects
with the kind of ownership systems developed in Ownership Types
\cite{simpleOwnership,existOwn} as used in Rust
\cite{RustPL2,RustBelt18,funk-ownership2024} and before that, in a number of approaches
to heap analyses \cite{PotterNC98,HillNP02,MitECOOP06} and
program verification
\cite{BoyLisShrPOPL03,hypervisor} including Spec$\#$
\cite{BarLeiSch05,BarDelFahLeiSch04} and universes
\cite{DieDroMue07,DietlMueller05,LuPotPOPL06},
and Borrowable Fractional Ownership \cite{borrow-fract-vmcai2024},
and recently integrated into OCAML \cite{ocaml-ownership-icfp2024}.


By enforcing encapsulation, \sd{all} % used to say both
these approaches share similarity with techniques such as
ownership types \cite{ownalias,NobPotVitECOOP98}, which also
protect internal implementation objects from accesses that cross
encapsulation boundaries.  Banerjee and Naumann demonstrated that by
ensuring confinement, ownership
systems can enforce representation independence (a property close to
``robust safety'') some time ago \cite{Banerjee:2005}.


The key point of Ownership is that the \textit{owned} internal
subcomponents of an object are accessible only via that object.
This
is quite closely related to our \textit{protected} capabilities,
however where protected capabilites are protected relative to some
protecting module, owned capabilities are owned with respect to all
modules in the system.  Conversely, we can characterise ownership as
protection taming \textit{all} references to an object: in practice,
that requires that the capability we protecting is unique at the time
the protection is established: \textit{ownership $==$ protection $+$
  uniqueness}.  This relationship has at least two potential avenues
for further work: first, that where this paper ensures protection via
a Hoare logic, we may be able to ensure protection by a
more straightforward type-based approach, by weakening some of the
guarantees provided by ownership types.  Second,  by tightening some
rules in our current Hoare logic (e.g.\ BLAH which establishes that a
capability is protected) we may be able to extend into a ``native''
Hoare logic of ownership.







