\documentclass[runningheads]{llncs}

\usepackage{ amssymb }
\usepackage{stmaryrd}

%\usepackage{subcaption} %kjx
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage{xspace, listings, lstcustom, wrapfig, graphicx, enumerate}
\usepackage{paralist}
\usepackage{color,colortbl, relsize}
\usepackage{rotating}
\usepackage{pifont}
\usepackage{multirow}
\usepackage{soul}
\usepackage{tcolorbox}
\usepackage[scaled=.9, light]{zlmtt}
\usepackage{siunitx}
\usepackage{setspace}
\usepackage{enumitem}

\include{macros}

\usepackage{times}
\usepackage{latexsym}
\usepackage{listings}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{ %
  language=Java,                % the language of the code
  mathescape=true,
  basicstyle=\footnotesize\tt,           % the size of the fonts that are used for the code
%  numbers=left,                   % where to put the line-numbers
%  numberstyle=\tiny\color{dkgreen},  % the style that is used for the line-numbers
%  stepnumber=1,                   % the step between two line-numbers. If it's 1, each line
                                  % will be numbered
%  numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},      % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
%  frame=single,                   % adds a frame around the code
  rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
  tabsize=2,                      % sets default tabsize to 2 spaces
  captionpos=b,                   % sets the caption-position to bottom
  breaklines=true,                % sets automatic line breaking
  breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  title=\lstname,                   % show the filename of files included with \lstinputlisting;
                                  % also try caption instead of title
  keywordstyle=\color{blue},          % keyword style
  commentstyle=\color{gray},       % comment style
  stringstyle=\color{mauve},         % string literal style
 % escapeinside={\%*}{*)},            % if you want to add LaTeX within your code
  morekeywords={PRE,POST,result,assume,method,mthd,function,fresh,assert,private,then,elseif,public,final,this,throw,new,||,to,def,any,fun,fld,abstract,policy,specification,ghost,field,func}        }  
         % if you want to add more keywords to the set
 


\newcommand{\kjx}[1]{{\color{orange}{KJX: #1}}}
\newcommand{\scd}[1]{{\color{dkgreen}{SD: #1}}}
\newcommand{\sdN}[1]{{\color{dkgreen}{#1}}}
\newcommand{\sdcomment}[1]{{\ensuremath{\blacksquare}}\footnote{\color{dkgreen}{SD: #1}}}
\newcommand{\secomment}[1]{{\ensuremath{\blacksquare}}\footnote{\se{#1}}}
\newcommand{\jncomment}[1]{{\ensuremath{\blacksquare}}\footnote{\kjx{#1}}}

 \newcommand{\sd}[1]{#1} % {{\color{dkgreen}{#1}}}% {#1} %
\newcommand{\tobyM}[1]{#1} %[1]{{\color{purple}{Toby: #1}}}
\newcommand{\se}[1]{#1} %{{\color{blue}{#1}}}


\newcommand{\ponders}[3]{\marginpar{\tiny\itshape\raggedright\textcolor{#2}{\textbf{#1:} #3}}\ignorespaces}
\marginparwidth=1.6cm \marginparsep=0cm
\newcommand{\TODO}[1]{{\color{red}#1}}
\newcommand{\sophia}[1]{} % {\ponders{Sophia}{dkgreen}{#1}}
\newcommand{\toby}[1]{} % {\ponders{Toby}{purple}{#1}}
\newcommand{\susan}[1]{} %{\ponders{Susan}{blue}{#1}}
\newcommand{\james}[1]{} % {\ponders{James}{orange}{#1}}

\begin{document}

\title{Holistic Specifications for Robust Programs}
%\titlerunning{Abbreviated paper title}

\author{Anonymous}


%\author{Sophia Drossopoulou\inst{1}\orcidID{0000-1111-2222-3333} \and
%   James Noble\inst{2}\orcidID{0000-1111-2222-3333} \and
%  Julian Mackay\inst{2}\orcidID{0000-1111-2222-3333} \and
% Susan Eisenbach\inst{1}\orcidID{0000-1111-2222-3333}}
%
%\authorrunning{S. Drossopoulou, J. Noble, et al.}
%% First names are abbreviated in the running head.
%% If there are more than two authors, 'et al.' is used.
%
%
%\institute{Imperial College London
%\email{\{scd,susan\}@springer.com} \and
%%\url{http://www.springer.com/gp/computer-science/lncs}
%Victoria University of Wellington
%\email{\{julian.mackay,kjx\}@ecs.vuw.ac.nz}}

\maketitle

\begin{abstract}
Functional specifications describe what
program components \emph{can} do: the \emph{sufficient} conditions to
invoke a component's operations.
They allow us to reason about the  use  of  
components in  the \emph{closed world} setting, where
the component interacts with known client code, and 
where the client code must establish the appropriate pre-conditions 
before calling into the component.
%
%%  a client who supplies arguments
%% meeting that operation's preconditions can invoke it, 
%%   and obtain the associated effect.

%don't start sentences with However!
Sufficient conditions are not enough to reason about 
the use of components in the \emph{open world} setting, where
the component interacts with external code, possibly of unknown
provenance, 
and where the  component itself may evolve over time. 
%
In this   open world  setting,
%%
we must also consider the
%% ensuring that your component is robust even when executing 
%% with buggy or malicious external code is critical.
%%  \emph{Holistic specifications}
%% --- as their name implies --- 
%% are concerned  with the \emph{overall} behaviour of a component,  in all possible 
%% interleavings of calls to the component's operations with those of the external code.
%% Thus, holistic specifications are concerned with \emph{sufficient} conditions, \ie
%% what is enough to \emph{cause} some effect, as well as with
%%
\emph{necessary} conditions, \ie
what are the conditions without which an effect will \emph{not} happen. 
%
In this paper we
%argue for the need for such holistic specifications,
 propose a 
 language \Chainmail for writing \emph{holistic} specifications that
 focus on necessary conditions (as well as sufficient conditions). We
 give a formal semantics for \Chainmail, and the core of \Chainmail has been
  mechanised in the Coq proof assistant.
%% By adopting holistic specification techniques like \Chainmail,
%% programmers can explicitly define what their components should not do,
%% making it easier to write
%% robust and reliable programs.
\end{abstract}



%% TODO: \\
%% Add equivalences of assertions, eg Wll(A) == Will(Will(A))\\
%% WIll(A) = A or Next(A or Will(A))

\section{Introduction}
\input{introductionWithWallet}

\section{Motivating Example: The Bank}
\label{sect:motivate:Bank}
\input{motivateBankShort}



\section{\Chainmail\ Overview}
\label{sect:chainmail}
\input{overChainmail}


\section{Overview of the Formal foundations}
\label{sect:formal}
We now give an overview of the formal model for \Chainmail. In section \ref{sect:overviewmodel} we 
introduce  the shape of the judgments used to give semantics to \Chainmail, while in section \ref{sect:PL} 
we describe the most salient aspects of an underlying programming language used in \Chainmail.

\subsection{\Chainmail\ judgments}
\label{sect:overviewmodel}
\input{overModel}

\subsection{An underlying programming language, \LangOO}
\label{sect:PL}
\input{summaryExecution}

\section{Assertions}
\label{sect:assertions}
\input{assertions}

%% \section{Examples}
%% \label{sect:example}

%% In this section we look at a couple of further examples from the
%% literature where a holistic specification would provide obvious
%% benefits in ensuring robustness.

%% \subsection{Attenuating the DOM}
%% \label{sect:example:DOM}
%% \input{DOM}  %%dunnow why this was ExampleDOM

%% \subsection{Authorizing the ERC20}
%% \label{sect:example:ERC20}
%% \sd{\input{ERC20}}


%% \subsection{Defending the DAO}

%% \sd{We specify the DAO in appendix \ref{Dao:appendix}}
%% %\input{DAO}

%% \section{Discussion}
%% \label{sect:discussion} 
%% \input{discussion}

\section{Examplar Driven Design}
\label{sect:discussion} 
\input{example-driven-design}

\section{Related Work}
\label{sect:related}
\input{related}

\section{Conclusion}
\label{sect:conclusion}
\input{conclusion}

\bibliographystyle{splncs04}
\bibliography{Case,more}


\newpage
\appendix

\section{Foundations: Programming Language}
\label{app:LangOO}
\input{langOO}

\section{Foundations: Specification Language}
\label{app:assertions}
\input{assertionsAppendix}

\section{Exemplar: Bank Account}
\label{Bank:appendix}
\input{Bank.appendix}

\section{Examplar: Authorising ERC20}
\label{sect:example:ERC20}
{\input{ERC20}}

\subsection{Example -- ERC20, the traditional specification }
\label{ERC20:appendix}
\input{ERC20.appendix.tex}

\section{Examplar: Defending the DAO}
\label{Dao:appendix}
\input{DAO}

\section{Examplar: Attenuating the DOM}
\label{sect:example:DOM}
\input{DOM}

\section{Coq Formalism}
\label{sect:coq}
\input{COQ}

%% \newpage
%% replaced for ACM-ization
%% % \small{
%%   \bibliographystyle{plain}
%% \onecolumn{
%%  \bibliography{Case,more}
%% }

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\forget{
\input{motivateDOM}

\input{DOM.appendix}




\section{Example -- DAO}
\input{DAO}

\section{Example -- Purse and Mint}
take from our earlier works

In two versions: one where there is a ledger inside the Mint, and one where the Mint has no path to the Purses. This will serve to demonstrate how \prg{internal} is supposed to work.

\section{Example -- Membrane}

TODO - take from ShuPeng's thesis
}

\include{rest}

\subsection*{LATEX mysteries and terminology}
 \begin{enumerate}
 \item
 How can we make the references refer to the Definitions, Lemmas etc
 rather than the section where these appear?
 \begin{quotation}
   \color{orange} KJX:   Not sure what the problem is. I've put labels
   in the definitions and I can use refs to get definition
   numbers~\ref{defONE} and~\ref{def:syntax:classes} ---
   not ~\ref{secONE} and ~\ref{sec:syntax:classes}, the section
   numbers containing those definions.

   Alternatively there is the ``cleverer'' package
   \url{http://tug.ctan.org/tex-archive/macros/latex/contrib/cleveref/cleveref.pdf}
   where a ``\verb+\cref{foo}+'' can generate both the type and the
   number e.g.  ``Definition 3''.
 \end{quotation}

 \item
 Need a nice metavariable for set of addresses, currently it is $R$. Perhaps instead use an enumeration, as eg $\{ \ \alpha_1,...\alpha_n\ \} $
 or $\kappa$?

 \begin{quotation}
 \color{orange} KJX: Hmm, the enumeration is fine. Otherwise $A$?
 $\mathcal{A}$?  Or we could call that set a ``footprint'' and so go
 with $F$ or $\mathcal{F}$\ldots
 \end{quotation}
\item
Find a nice term  to refer to module pairs  (internal, external), and a term for
our version visible states semantics.
 \begin{quotation}
   \color{orange} KJX: ``modules'' and ``modular state semantics''.
   Going to ``modules'' only makes sense with my answer below.
   Other permutations of
   ``visible module/modular state semantics'' work also work:
     modular visible state; visible modular state; etc\ldots
 \end{quotation}


\item
Better symbols for module linking (currently a $\M\link\M'$), and
for module pairing (currently a $\M\mkpair \M'$) -- perhaps there should not be such an operator, as
it does not create a new module, it is only used in execution ($\M\mkpair \M', \sigma \leadsto \sigma'$)
and in satisfaction of assertions ($\M\mkpair \M', \sigma\models \A$).
\footnote{\toby{TM: I like~$\M \mkpair \M'$ as it suggests the asymmetry of the visible
    state semantics wrt~$\M$ and~$\M'$.}}

 \begin{quotation}
   \color{orange} KJX: I'm so used to $\M\link\M'$ that I can't think
   of an alternative --- or do I recall we used $\M * \M'$ as a
   separating conjunction?

   So I really liked $\M\mkpair \M'$ --- except then I though that I
   couldn't remember which was the module (inside) and which the
   anti-module (the outside).  For some reason I thought outside would
   go first.  Then I realised, it's easy, cos $\M$ is always the
   module, and $\M'$ is the antimodule.

   At which point I though: OK so let's just write $\M$ as the module,
   and given any $\M$, then $\M'$ (or $\overline{\M}$ or I guess
   \textbf{out}($\M$)) for the antimodule.

   The only thing I think this loses is that the $\M\mkpair \M'$
   syntax, also like a separating conjunction, is sort of
   self-framing: $\M\mkpair \M'$ encompasses the universe of modules.
   Whereas the other way around, we'd need a (implicit) universe of
   all modules $\mathcal{U}$, and then define $\M' \triangleq
   \mathcal{U} - \M$  If we went with $\M'$ then Sophia couldn't use
   $\M''$ and friends --- have to write \texttt{N} and \texttt{O} for other
   modules?

   I think the only change I could see in the whole document was that
   lemma~\ref{lemma:module_pair_execution} is subsumed into
   lemma~\ref{lemma:linking:properties}.
 \end{quotation}


 \end{enumerate}
