\section{Appendix to Section \ref{sect:underlying} -- The programming language \LangOO}
\label{app:loo}


We introduce \LangOO, a simple, typed, class-based, object-oriented language.

\subsection{Syntax}

The syntax of \LangOO is given in Fig. \ref{f:loo-syntax}\footnote{{Our motivating example is provided in a slightly richer syntax for greater readability.}}.
To reduce the complexity of our formal models, as is usually done, CITE - CITE,  \LangOO lacks many
common languages features, omitting static fields and methods, interfaces,
inheritance, subsumption, exceptions, and control flow.  
 \LangOO % includes ghost fields,  that may only be used in the specification language.
and which may be defined recursively.

\LangOO modules ($M$) map class names ($C$) to class definitions ($\textit{ClassDef}$).
A class definition consists of % \jm[]{an optional annotation \enclosed},
a list of field definitions, ghost field definitions, and method definitions.
{Fields, ghost fields, and methods all have types, $C$; {types are
    classes}.
    Ghost fields may be optionally 
 annotated as \texttt{intrnl}, requiring the argument to have an internal type, and the 
body of the ghost field to only contain references to internal objects. This is enforced by the
limited type system of \LangOO.}
A program state ($\sigma$) is a pair of of a stack and a heap.
The stack is a a stack is a non-empty list of frames ($\phi$), and the heal ($\chi$)
is a map from addresses ($\alpha$) to objects ($o$). A frame consists of a local variable
map and a continuation .\prg{cont} that represents the statements that are yet to be executed ($s$).
% or a hole waiting to be filled by a method return in the frame above ($x := \bullet; s$).
A statement is either a field read ($x := y.f$), a field write ($x.f := y$), a method call
($u :=y_0.m(\overline{y})$), a constructor call ($\prg{new}\ C$), 
% a method return statement ($\prg{return}\ x$), 
  a sequence of statements ($s;\ s$),
  or empty ($\epsilon$).


\LangOO also includes syntax for expressions $\re$ that may %only
be used in writing
specifications or the definition of ghost fields.





\subsection{Semantics}
\LangOO is a simple object oriented language, and the operational semantics 
(given in Fig. \ref{f:loo-semantics} and discussed later)
do not introduce any novel or surprising features. The operational 
semantics make use of several helper definitions that we 
define here.

{
We provide a definition of reference interpretation in Definition \ref{def:interpret}
\begin{definition}
\label{def:interpret}
For a frame $\phi= (\overline {x \mapsto v}, s)$, and a program state $\sigma = (\overline \phi \cdot \phi,, \chi)$, we   define:
\begin{itemize}
\item
$\interpret{\phi}{x}\ \triangleq\ v_i$\ \ \ if \ \ \ $x=x_i$
\item
 $\interpret{\sigma}{x}\ \triangleq\  \interpret{\phi}{x}$
\item
$\interpret{\sigma}{\alpha.f}\ \triangleq\ v_i $ \ \ if \ \ $\chi(\alpha)=(\_; \  \overline {f \mapsto v})$, and $f_i=f$
\item
$\interpret{\sigma}{x.f}\ \triangleq\ \interpret{\sigma}{\alpha.f}$ where $\interpret{\sigma}{x}=\alpha$
\item
$\phi.\prg{cont} \ \triangleq\ s$ 
\item
$\sigma.\prg{cont} \ \triangleq\ \phi.\prg{cont}$\
\item
$\phi[\prg{cont}\mapsto s'] \ \triangleq\ (\overline {x \mapsto v}, s')$
\item
$\sigma[\prg{cont}\mapsto s'] \ \triangleq \ (\ {\overline \phi}\cdot \phi[\prg{cont}\mapsto s'],\  \chi\ )$ 
\item
$\phi[\prg{x'}\mapsto v'] \ \triangleq\ ( \ (\overline {x \mapsto v})[\prg{x'}\mapsto v'],\ s \ )$
\item
$\sigma[\prg{x'}\mapsto v'] \ \triangleq\ (\ (\overline {\phi} \cdot (\phi[\prg{x'}\mapsto v']), \ \chi)$ 
\item
$\sigma [\alpha \mapsto o ] \ \triangleq\ (\ (\overline {\phi} \cdot \phi), \ \chi [\alpha \mapsto o ]\ )$ 
\item
$\sigma [\alpha.f' \mapsto v' ] \ \triangleq\ \sigma [\alpha \mapsto o ] $\ \ \  if \ \  
$\chi(\alpha)=(C, {\overline {f \mapsto v}})$, and $o=(\ C;  ({\overline {f \mapsto v}})[f' \mapsto v' ]\ )$ 
\end{itemize}
\end{definition}
}
That is, a variable $x$, or a field access on a variable $x.f$ 
has an interpretation within a program state of value $v$
if $x$ maps to $v$ in the local variable map, or the field
$f$ of the object identified by $x$ points to $v$.

Definition \ref{def:class-lookup} defines the class lookup function an object 
identified by variable $x$.
\begin{definition}[Class Lookup]
\label{def:class-lookup}
For program state $\sigma = ({\overline {\phi}}\cdot\phi, \chi)$, class lookup is defined as 
$$\class{\sigma}{x}\ \triangleq\ C \ \ \ \ \ \mbox{if} \ \ \  \chi(\interpret{\sigma}{x})=(C,\_ )$$
\end{definition}

Module linking is defined for modules with disjoint definitions:

\begin{definition}
\label{def:linking}
For all modules $\Mtwo$ and $M$, if the domains of $\Mtwo$ and $M$ are disjoint, 
we define the module linking function as $M\cdot \Mtwo\ \triangleq\ M\ \cup\ M'$.
\end{definition}
That is,  their linking is the union of the two if their domains are disjoint.

Definition \ref{def:meth-lookup} defines the method lookup function for a method
call $m$ on an object of class $C$.
\begin{definition}[Method Lookup]
\label{def:meth-lookup}
For module $\Mtwo$, class $C$, and method name $m$, method lookup is defined as 
$$\meth{\Mtwo}{C}{m}\ \triangleq\ { pr}\  \prg{method}\ m\ (\overline{x : T}) {:T}\{\ s\ \}  $$
if there exists an $M$ in $\Mtwo$, so that $M(C)$ contains the definition ${ pr}\  \prg{method}\ m\ (\overline{x : T}) {:T}\{\ s\ \} $
\end{definition}

Definition \ref{def:fields-lookup} looks up all the field identifiers in a given class
\begin{definition}[Fields Lookup]
\label{def:fields-lookup}
For modules $\Mtwo$,and  class $C$, fields lookup is defined as 
$$fields(\Mtwo,C) \ \triangleq\  \  \{ \ f  \ | \  \exists  M\in\Mtwo. s.t.  M(C) \mbox{contains the definition}  \prg{field}\ f: T\ \} $$
\end{definition}



We define what it means for two objects to come from the same module
\begin{definition}[Same Module]
\label{def:same:module}
For program state $\sigma$,  modules $\Mtwo$, and variables $x$ and $y$, we defone
$$\Same {x} {y} {\sigma}{\Mtwo}\ \triangleq\ \exists C, C', M[ \ M\in \Mtwo \wedge C, C'\in M \wedge  \class{\sigma}{x}=C \wedge \class{\sigma}{y} =C'\ ]$$
\end{definition}

As we already said in \S \ref{s:underlying}, we forbid assignments to a method's parameters. 
To do that, the following function returns the  identifiers of the formal parameters of the currently active method.



\begin{definition}
For program state $\sigma$:
\label{def:params}

$\Formals \sigma \Mtwo \ \ \triangleq \ \  \overline x \ \ \ \mbox{such that} \ \  \  \exists \overline {\phi},\,\phi_k, \, \phi_{k+1}, \,C,\,p.$\\
$\strut \hspace{3.2cm} [   \ \ \sigma =  (\overline {\phi}\cdot{\phi_k}\cdot{\phi_{k+1}}\,, \chi) 
\  \ \wedge\  \ \phi_k.\prg{cont}=\_:=y_0.m(\_); \_ \  \  \wedge\ \ $\\
$\strut \hspace{3.2cm} \class{(\phi_{k+1},\chi)}  {\prg{this} }\  \ \wedge\ \ \meth{\Mtwo} {C} {m} = p \ C::m(\overline{x : \_}){:\_}\{\_\} \ ] $
\end{definition}





While the small-step operational semantics of \LangOO is given in Fig. \ref{f:loo-semantics},
specification satisfaction is defined over an abstracted notion of 
the operational semantics that models the open world. %, called \jm[]{\emph{external states semantics}}. 




An \emph{Initial} program state contains a single frame 
with a single local variable \prg{this} pointing to a single object 
in the heap of class \prg{Object}, and a continuation.
\begin{definition}[Initial Program State]
\label{def:initial}
A program state $\sigma$ is said to be an initial state ($\initial{\sigma}$)
if and only if
\begin{itemize}
\item
$\sigma\ =\  ( ((\prg{this}\ \mapsto\ \alpha), s); \  (\alpha \mapsto (\prg{Object}, \emptyset)$
\end{itemize} 
for some address $\alpha$ and some statement $s$.
\end{definition}


%We give the semantics of module pair execution in Definition \ref{def:pair-reduce}
%\begin{definition}[External State Semantics]
%\label{def:pair-reduce-appendix}
%For all internal modules $M_1$, external modules $M_2$, and program configurations $\sigma$ and $\sigma'$, 
%we say that $\reduction{M_1}{M_2}{\sigma}{\sigma'}$ if and only if
%\begin{itemize}
%\item
%$\class{\sigma}{\sigma(\prg{this})}\ \in\ M_2$ and
%\item
%$\class{\sigma'}{\sigma'(\prg{this})}\ \in\ M_2$ and 
%\end{itemize} 
%and
%\begin{itemize}
%\item
%$\exec{M_1\ \circ\ M_2}{\sigma}{\sigma'}$ or
%\item
%$M_1 \circ M_2, \sigma \leadsto \sigma_1 \leadsto \ldots \sigma_n \leadsto \sigma'$ and $\class{\sigma_i}{\sigma_i(\prg{this})} \in M_1$ for all $1 \leq i \leq\ n$
%\end{itemize}
%\end{definition}

We provide a semantics for expression evaluation is given in Fig. \ref{f:evaluation}. 
That is, given a module $M$ and a program state $\sigma$, expression $e$ evaluates to $v$
if $\eval{M}{\sigma}{e}{v}$. Note, the evaluation of expressions is separate from the operational
semantics of \LangOO, and thus there is no restriction on field access.

{
\paragraph{Lemmas and Proofs}
}
%We prove lemma  \ref{l:var:unaffect}, using the following lemma:

%\begin{lemma}
%\label{l:leadsto:depth}
%For any states $\sigma$, $\sigma'$, modules $\Mtwo$,  number $k$, and variable $y$:
%
%\begin{enumerate}
%\item
%$ \EarlierS {\sigma}  {\sigma'}\ \ \Longrightarrow \ \ \DepthSt {\sigma} \leq \DepthSt {\sigma}$.
%
%\item
%$\leadstoBoundedThree   {\Mtwo}  {\sigma_1} {\sigma}  {\sigma_2}\ \  \ \Longrightarrow\ \ \ \DepthSt {\sigma} \leq \DepthSt {\sigma_1}\ \wedge \ \DepthSt {\sigma} \leq \DepthSt {\sigma_2}$.
%
%\item
%$\leadstoBoundedThree   {\Mtwo}  {\sigma_1} {\sigma}  {\sigma_2}\ \wedge\ 
% k=    \DepthSt \sigma \ \wedge\ (k <\DepthSt {\sigma_1} \vee k < \DepthSt {\sigma_2})\ \ \ \Longrightarrow \ \ \
%\interpret {\sigma} {y} = \interpret {\RestictTo {\sigma_1} {k}} y= \interpret {\RestictTo {\sigma_2} {k}} y .$
%
%\item
%$\leadstoOrig {\Mtwo} {\sigma} {\sigma'}\ \  \wedge\  \  \DepthSt \sigma = \DepthSt {\sigma'} \ \ \wedge \ \
%y\notin \vs(\sigma.\prg{cont}) \ \ \Longrightarrow \ \  \interpret \sigma y =  \interpret {\sigma'} y$
% 
%\end{enumerate}
%\end{lemma}
%}
% 
%{
%\beginProof{l:leadsto:depth}
%\begin{enumerate}
%\item Follows from the definition of $\DepthSt {\_}$, and $ \EarlierS {\_}  {\_}$.
%\item Follows from the definition  $\leadstoBoundedThree {\_} {\_} {\_} {\_}$ and (1).
% 
%\item  From. $\leadstoBoundedThree   {\Mtwo}  {\sigma_1} {\sigma}  {\sigma_2}\ \wedge\ 
% k= \DepthSt \sigma \ \wedge\ (k <\DepthSt {\sigma_1} \vee k < \DepthSt {\sigma_2})$ we can deduce that the step from $\sigma_1$ to $\sigma_2$ 
% is either a method call from $\sigma$, 
% ????
% 
%\item
%Follows from the operational semantics
%
%\end{enumerate}
%\completeProof
%}

 
\beginProof{l:wf:state}
{The first assertion is proven by unfolding the definition of $\_ \models \_ $.

The second assertion is proven by case analysis on the execution relation $\exec {\_} {\sigma} {\sigma'}$. 
The assertion gets established when we call a method, and is preserved through all the execution steps, because we do not allow assignments to the formal parameters.
 
}
\completeProof

We now prove lemma \ref{l:var:unaffect}:

\beginProof{l:var:unaffect} 
\begin{itemize}
\item
We first show that $\leadstoBoundedThree {\Mtwo} {\sigma} {\sigma\bd}  {\sigma'}   \ \wedge \ k<\DepthSt \sigma\bd  \ \ \Longrightarrow \ \  \interpret {\RestrictTo \sigma k} y =  \interpret {\RestrictTo {\sigma'} k} y$
This follows easily from the operational semantics, and the definitions.


\item
By induction on the earlier part, we  obtain that $\leadstoBoundedStar {\Mtwo}  {\sigma}  {\sigma'}  \ \wedge \ k<\DepthSt \sigma  \ \ \Longrightarrow \ \  \interpret {\RestrictTo \sigma k} y =  \interpret {\RestrictTo {\sigma'} k} y$

\item

We now show that $\leadstoBoundedStarFin {\Mtwo}  {\sigma}  {\sigma'} \ \wedge \ y\notin \vs(\sigma.\prg{cont}) \ \ \Longrightarrow \ \  \interpret \sigma y =  \interpret {\sigma'} y$ by  induction on the number of steps, and  using the earlier lemma.
\end{itemize}
\completeProof

{Lemma \ref{lemma:relevant:more} states that initila states are well-formed, and that (\ref{threeLR}) a pre-existing object, locally reachable after any number of scoped execution steps, was locally reachable at the first step.


\begin{lemma}
\label{lemma:relevant:more}
For all modules $\Mtwo$, states $\sigma$, $\sigma'$,   and frame $\phi$:
\begin{enumerate}
\item
\label{lemma:relevant:more:one}
$\initial {\sigma}  \ \    \Longrightarrow \ \ \Mtwo \models \sigma $
\item
\label{threeLR}
{${\leadstoBoundedStar {\Mtwo}  {\sigma}    {\sigma'}}  \ \  \Longrightarrow\ \ 
dom(\sigma) \cap \LRelevantO {\sigma'} \subseteq   \LRelevantO {\sigma}$
}

\end{enumerate}
\end{lemma}

{Consider Fig.  \ref{fig:illusrPreserve} . %, and Fig.  \ref{fig:UpSemanticsBounded}.
Lemma \ref{lemma:relevant:more}, part \ref{threeLR}  promises that any objects locally reachable in $\sigma_{14}$ which already existed in $\sigma_{8}$, were locally reachable in $\sigma_{8}$. However, the lemma is only  applicable to scoped execution, and as 
$\notLeadstoBoundedStar {\Mtwo} {\sigma_8}  {\sigma_{17}}$, 
the lemma does not promise that  objects locally reachable in $\sigma_{17}$ which already existed in $\sigma_{8}$, were locally accessible in $\sigma_{8}$ -- namely it could be that objects are made globally reachable upon method return, during the step from $\sigma_{14}$ to $\sigma_{15}$.}
}

\vspace{1cm}
Finally, we define the evaluation of expressions, which, as we already said, represent ghost code.




\begin{figure}[hbp]
\begin{minipage}{\textwidth}
\footnotesize{
\begin{mathpar}
\infer
		{}
		{\eval{M}{\sigma}{v}{v}}
		\quad(\textsc{E-Val})
		\and
\infer
		{}
		{\eval{M}{\sigma}{x}{\interpret{\sigma}{x}}}
		\quad(\textsc{E-Var})
		\and
%\infer
%		{
%		\eval{M}{\sigma}{e_1}{i_1}\\
%		\eval{M}{\sigma}{e_2}{i_2}\\
%		i_1 + i_2 = i
%		}
%		{
%		\eval{M}{\sigma}{e_1 + e_2}{i}
%		}
%		\quad(\textsc{E-Add})
%		\and
%\infer
%		{
%		\eval{M}{\sigma}{e_1}{v}\\
%		\eval{M}{\sigma}{e_2}{v}
%		}
%		{
%		\eval{M}{\sigma}{e_1 = e_2}{\true}
%		}
%		\quad(\textsc{E-Eq}_1)
%		\and
%\infer
%		{
%		\eval{M}{\sigma}{e_1}{v_1}\\
%		\eval{M}{\sigma}{e_2}{v_2}\\
%		v_1 \neq\ v_2
%		}
%		{
%		\eval{M}{\sigma}{e_1 = e_2}{\false}
%		}
%		\quad(\textsc{E-Eq}_2)
%		\and
%\infer
%		{
%		\eval{M}{\sigma}{e}{\true}\\
%		\eval{M}{\sigma}{e_1}{v}
%		}
%		{
%%		\eval{M}{\sigma}{\ifthenelse{e}{e_1}{e_2}}{v}
%		\eval{M}{\sigma}{e}{v}
%		}
%		\quad(\textsc{E-If}_1)
%		\and
%\infer
%		{
%		\eval{M}{\sigma}{e}{\false}\\
%		\eval{M}{\sigma}{e_2}{v}
%		}
%		{
%%		\eval{M}{\sigma}{\ifthenelse{e}{e_1}{e_2}}{v}
%		\eval{M}{\sigma}{e}{v}
%		}
%		\quad(\textsc{E-If}_2)
%		\and
\infer
		{
		\eval{M}{\sigma}{\re}{\alpha}
		}
		{
		\eval{M}{\sigma}{\re.f}{\interpret{\sigma}{\alpha.f}}
		}
		\quad(\textsc{E-Field})
		\and
\infer
		{
		\eval{M}{\sigma}{\re_0}{\alpha}\\
		\overline {\eval{M}{\sigma}{\re}{v}}\\
		M(\class{\sigma}{\alpha})\ \mbox{contains}\ \prg{ghost}\ gf(\overline{x : T)}\{\re\} : T' \\
		\eval{M}{\sigma}{[\overline {v/x}]\re}{v}
		}
		{
		\eval{M}{\sigma}{\re_0.gf(\overline{\re})}{v}
		}
		\quad(\textsc{E-Ghost})
\end{mathpar}
}
\caption{\LangOO Expression evaluation}
\label{f:evaluation}
\end{minipage}
\end{figure}

